Dark energy is based on a math error from 1930

### 2024-12-27 paper/paper.tex
Halfway through the derivation...
... but it's past the kiddo's bed time.

The measurement of Type Ia supernovae is one of the primary sources of data for
cosmological models. The measurements involve estimating the peak magnitude in
the B filter \citet{riess1998} for the rest frame, meaning we want to know how
bright the SnIa would be if its light didn't undergo redshift. Redshift causes
the light that would be mesaured by the B filter in a rest frame to be shifted
to another filter. In order to observe a SnIa in an arbitrary filter and report
the magnitude in the B filter we use a technique called k-corrections.

The first formal derivation of k-corrections was performed by
\citet{tollman1930}, and it failed to account for bandwidth stretching, one of
the dimming effects of redshift. This error was noted by \citet{desitter1934},
but in \citet{hubble1935} the descrepancy was noted and ignored.
\citet{oke1968} rederived k-corrections, but they failed to account for time
dilation, one of the other dimming effects of redshift. In \citet{kim1996},
k-corrections were extended to handle additional cross-filter comparisons and
address zero-point corrections, but they started their derivation by
referencing the incorrect equation from \citet{oke1968}. \citet{hsiao2007}
expanded on the work done by \citet{kim1996} by providing expected SnIa
template spectra. Finally, \citet{burns2010} implemented the popular software
package Sn(OO)py which implemented the SnIa template spectra into
k-corrections. This software package computes k-corrections differently from
prior work, and in order to match prior values, the code adds an extra term.
This history is explored in more depth in Section \ref{sec:history}.

The effect of using an incorrect magnitude for SnIa is that we think objects
are farther away than they really are, and the effect componds for greater
distances. Up until \citet{riess1998}, we didn't have distant enough
observations for this error to matter much for cosmology models, but in the
late 1990s, it was clear that our measurements for distance and redshift were
not linear. This observation led to a model that utilized a cosmological
constant and dark energy in order to explain the non-linear distance-redshift
graph. We will show in Section \ref{sec:correction} that correcting the flaw
with k-corrections leads to a linear graph that does not need to rely on a
cosmological constant.

Since \citet{planck2020}, it has been evident that something was missing with
our understanding of cosmology. This research used an alternative technique
based on measurements of the cosmic microwave background to measure the Hubble
constant, but as summarized in \citet{perivolaropoulos2022}, this measurement
was incompatable with the Hubble constant measured from SnIa. As we will
discuss in Section \ref{sec:tension}, fixing the error with k-corrections
produces a measurement of the Hubble constant that is compatible with models
based on the cosmic microwave background.

\section{A brief history of k-corrections}
\label{sec:history}

@@

Measuring the apperant brightness of a Type Ia supernova is non-trivial. The
aspect of the process that we will explore here concerns how redshift affects
the light we observe. This problem is often referred to as k-corrections, and
one of the first mathematical treatments of the problem was performed by
\citet{tolman1930}. However, when Tolman made his derivation, he did not
consider the effects of a spectra that is stretched out due to redshift.

@@

A year later, \citet{hubble1935} published a similar set of calculations for
k-corrections, but these equations used $(1 + z)^2$ instead of the $(1 + z)^3$
correction term used by de Sitter. They started their derivation by copying the
incorrect equation from 1930. After their derivation, they noted,

@@

The Hubble k-corrections with the incorrect correction term have been used ever
since.

By \citet{oke1968}, the two factors of $(1 + z)$ were attributed to the change
in energy and to the spectral bandwidth elongation, which leaves time dilation
as the factor that was omitted. A graph that demonstrates why it's essential to
correct for both the spectra bandwidth warping and time dilation is presented
in Figure \ref{fit:k-example}.

\begin{figure}[h]
  \includegraphics[width=\linewidth]{k-corrections_for_photon_counts.png}
  \caption{An example of how k-corrections take an observed magnitude and
  produce a rest frame magnitude. In this example, $z=1$. The observation
  filter measures magnitude, which is equivalent to measuring the number of
  photons per nanometer. In the bottom left panel, a correction of $1+z$ is
  applied which doubles the photons per nanometer. The top right panel shows
  the effect of correcting for the bandwidth spectrum warping effect --
  correcting all wavelengths for the rest frame means that the measured
  wavelengths have blueshifted, ideally into the cross band B filter. Applying
  both effects, shown in the bottom right panel, requires applying two
  correction factors of $1+z$. This example omits the correction for a similar
  effect related to lower energy due to the Planck relation because CCD cameras
  obviate the need to correct for this effect.
  }
  \label{fig:k-example}
\end{figure}

The modern treatment of k-corrections is based on the work of \citet{kim1996}.
This work extended the calculations of k-corrections to extend to filters
beyond B and V. It also introduced a term that deals with the zero-point for the
actual filters. Equation 2 from this paper is the form of the k-corrections
used in the modern day:

\begin{equation}
  K_{xy} = -2.5\text{log} \left(\frac{\int \mathcal{Z}(\lambda)S_x(\lambda)d\lambda}{\int \mathcal{Z}(\lambda)S_y(\lambda)d\lambda}\right) + 2.5\text{log}(1+z) + 2.5\text{log}\left(\frac{F(\lambda)S_x(\lambda)d\lambda}{F(1/(1+z))S_y(\lambda))d\lambda}\right).
\end{equation}

The first term of this equation deals with zero-points, which is a magnitude
added to individual filters to correct for the light gathering abilities of the
telescope. The third term measures the flux in the observation filter and divides it by  the flux that would be measured

The next major advance in k-corrections was

@@

\begin{minipage}[c]{0.95\textwidth}
\lstinputlisting[language=Python, caption=A demonstration that k-corrections
calculated by snpy are off by \texttt{2.5*log10(1+z)}. The roundoff error is
roughly \texttt{2e-16}.]{code_listing.py}
\end{minipage}

>>>>

The measurement of Type Ia supernovae is one of the primary sources of data for
cosmological models. The measurements involve estimating the peak magnitude in
the B filter \citet{riess1998} for the rest frame, meaning we want to know how
bright the SnIa would be if its light didn't undergo redshift. Redshift causes
the light that would be mesaured by the B filter in a rest frame to be shifted
to another filter. In order to observe a SnIa in an arbitrary filter and report
the magnitude in the B filter we use a technique called K-corrections.

The first formal derivation of K-corrections was performed by
\citet{tolman1930}, and it failed to account for bandwidth stretching, one of
the dimming effects of redshift. This error was noted by \citet{desitter1934},
but in \citet{hubble1935} the descrepancy was noted and ignored.
\citet{oke1968} rederived K-corrections, but they failed to account for time
dilation, one of the other dimming effects of redshift. In \citet{kim1996},
K-corrections were extended to handle additional cross-filter comparisons and
address zero-point corrections, but they started their derivation by
referencing the incorrect equation from \citet{oke1968}. \citet{hsiao2007}
expanded on the work done by \citet{kim1996} by providing expected SnIa
template spectra. Finally, \citet{burns2010} implemented the popular software
package Sn(OO)py which implemented the SnIa template spectra into
K-corrections. This software package computes K-corrections differently from
prior work, and in order to match prior values, the code adds an extra term.
This history is explored in more depth in Section \ref{sec:history}.

The effect of using an incorrect magnitude for SnIa is that we think objects
are farther away than they really are, and the effect componds for greater
distances. Up until \citet{riess1998}, we didn't have distant enough
observations for this error to matter much for cosmology models, but in the
late 1990s, it was clear that our measurements for distance and redshift were
not linear. This observation led to a model that utilized a cosmological
constant and dark energy in order to explain the non-linear distance-redshift
graph. We will show in Section \ref{sec:correction} that correcting the flaw
with K-corrections leads to a linear graph that does not need to rely on a
cosmological constant.

Since \citet{planck2020}, it has been evident that something was missing with
our understanding of cosmology. This research used an alternative technique
based on measurements of the cosmic microwave background to measure the Hubble
constant, but as summarized in \citet{perivolaropoulos2022}, this measurement
was incompatable with the Hubble constant measured from SnIa. As we will
discuss in Section \ref{sec:tension}, fixing the error with K-corrections
produces a measurement of the Hubble constant that is compatible with models
based on the cosmic microwave background.

\section{A brief history of K-corrections}
\label{sec:history}

@@

Measuring the apperant brightness of a Type Ia supernova is non-trivial. The
aspect of the process that we will explore here concerns how redshift affects
the light we observe. This problem is often referred to as K-corrections, and
one of the first mathematical treatments of the problem was performed by
\citet{tolman1930}. However, when Tolman made his derivation, he did not
consider the effects of a spectra that is stretched out due to redshift.

@@

A year later, \citet{hubble1935} published a similar set of calculations for
K-corrections, but these equations used $(1 + z)^2$ instead of the $(1 + z)^3$
correction term used by de Sitter. They started their derivation by copying the
incorrect equation from 1930. After their derivation, they noted,

@@

The Hubble K-corrections with the incorrect correction term have been used ever
since.

By \citet{oke1968}, the two factors of $(1 + z)$ were attributed to the change
in energy and to the spectral bandwidth elongation, which leaves time dilation
as the factor that was omitted. A graph that demonstrates why it's essential to
correct for both the spectra bandwidth warping and time dilation is presented
in Figure \ref{fig:k-example}.

\begin{figure}[h]
  \includegraphics[width=\linewidth]{K-corrections_for_photon_counts.png}
  \caption{An example of how K-corrections take an observed magnitude and
  produce a rest frame magnitude. In this example, $z=1$. The observation
  filter measures magnitude, which is equivalent to measuring the number of
  photons per nanometer. In the bottom left panel, a correction of $1+z$ is
  applied which doubles the photons per nanometer. The top right panel shows
  the effect of correcting for the bandwidth spectrum warping effect --
  correcting all wavelengths for the rest frame means that the measured
  wavelengths have blueshifted, ideally into the cross band B filter. Applying
  both effects, shown in the bottom right panel, requires applying two
  correction factors of $1+z$. This example omits the correction for a similar
  effect related to lower energy due to the Planck relation because CCD cameras
  obviate the need to correct for this effect.
  }
  \label{fig:k-example}
\end{figure}

The modern treatment of K-corrections is based on the work of \citet{kim1996}.
This work extended the calculations of K-corrections to extend to filters
beyond B and V. It also introduced a term that deals with the zero-point for the
actual filters. Equation 2 from this paper is the form of the K-corrections
used in the modern day:

\begin{equation}
\label{eq:kim2}
  K_{xy} =
    -2.5\text{log} \left(
      \frac{\int \mathcal{Z}(\lambda)S_x(\lambda)d\lambda}
           {\int \mathcal{Z}(\lambda)S_y(\lambda)d\lambda}\right)
    + 2.5\text{log}(1+z)
    + 2.5\text{log}\left(
      \frac{\int F(\lambda)S_x(\lambda)d\lambda}
           {\int F(\lambda/(1+z))S_y(\lambda))d\lambda}\right).
\end{equation}

This equation has two problems with it, which we will explore in Section
\ref{sec:derivation}.

\section{Derivation of K-corrections}
\label{sec:derivation}

The definition of K-corrections is

\begin{equation}
\begin{aligned}
\label{eq:definition}
  m_x &= M_x + \mu + K_{xy} \\
      &= M_x + m_y - M_y + K_{xy} \\
      &= m_y + K_{xy}.
\end{aligned}
\end{equation}

The second line of Equation \ref{eq:definition} expands the distance modulus
$\mu$ using $\mu = m = M$ where $m$ is the observed magnitude and $M$ is the
absolute magnitude. Since the absolute magnitude is an intrinsic property of a
stellar object, we can use $M = M_x = M_y$, so the absolute magnitude terms
cancel out in the third line.

We will also need to use the flux to magnitude formula:

\begin{equation}
\begin{aligned}
\label{eq:flux2mag}
  m_x &= -2.5 \text{log}(\mathcal{F}_x) + C_x \\
      &= -2.5 \text{log}(C'_x \mathcal{F}_x) .
\end{aligned}
\end{equation}

Here, $C_x$ represents the zero-point for the filter $x$ on some particular
telescope. In order to use consistend magnitude values across different
telescopes that have different light gathering powers, we take the measured
magnitude and multiply it by the ratio of the flux rate for the standard
telescope and filter to the flux rate for this particular telescope and
filter.

Since the K-correction is a magnitude value, we can define

\begin{equation}
K_x = -2.5\text{log}(k_x C_x)
\end{equation}

and use Equation \ref{eq:flux2mag} to recombine the flux term $\mathcal{F}_y$
with $m_y$ from Equation \ref{eq:definition}. This gives

\begin{equation}
\begin{aligned}
            m_x &= -2.5 \text{log}(\mathcal{F}_y) + K_{xy} \\
                &= -2.5 \left( \text{log}(\mathcal{F}_y) + \text{log}(k_x C_x) \right) \\
                &= -2.5 \text{log}(\mathcal{F}_y \times k_{xy} \times C_x) \\
  \mathcal{F}_x &= \mathcal{F}_y \times k_{xy} \times C_x .
\end{aligned}
\end{equation}

The work up to this point has been to translate the K-corrections equation from
a magnitude problem to a flux problem. This has intuitive benefits, at least
for the authors. The goal is to make $k_{xy}$ equal the ratio of the flux rate
that we want in filter $x$ to the flux rate that we observe in filter $y$.

The first term of this equation deals with zero-points $C$, which is a
magnitude added to individual filters to correct for the light gathering
abilities of the filter together with the telescope. We are converting the
magnitude from the $y$ filter to the $x$ filter by using $m_x = m_y - k_{xy}$,
so $k_{xy}$ needs to include $C_y - C_x$.

The numerator of the third term, $\int F(\lambda)S_x(\lambda)d\lambda$,
measures what the flux would be if there was no redshift in the target filter.
The denominator $\int F(\lambda/(1+z))S_y(\lambda))d\lambda$ takes the
observation filter's sensitivity $S_y(\lambda)$ at each wavelength $\lambda$
and multiples it by the rest frame spectral energy density $F(\lambda/(1+z))$
that corresponds to the blueshifted wavelength $\lambda / (1+z)$.

The effect of
multiplying the filter sensitivity $S(\lambda)$ by the blueshifted spectral energy density $F(\lambda / (1+z))$ is the same as if we created a new
spectral energy density $F'(\lambda) = F(\lambda / (1+z))$. The key thing to
notice with $F'$ is that it artificially inflates the total amount of flux by a
factor of $1 + z$. That is,

\begin{equation}
(1+z) \times \int_{a}^{b} S(\lambda) d\lambda = \int_{a(1+z)}^{b(1+z)} S'(\lambda) d\lambda .
\end{equation}

In order to discuss flux instead of magnitude, we can rewrite Equation \ref{eq:kim2} as

\begin{equation}
\begin{aligned}
\label{eq:kim_flux}
  K_{xy} &= -2.5 \text{log}(\mathcal{F}) + C_x - C_y \\
         &= -2.5 \text{log}\left(
              \frac{(1+z) \int F(\lambda)S_x(\lambda)d\lambda}
                   {\int F(1/(1+z))S_y(\lambda))d\lambda}\right) + C_x - C_y \\
  \mathcal{R} &= (1+z) \times \frac{\int F(\lambda)S_x(\lambda)d\lambda}
                                   {\int F(\lambda/(1+z))S_y(\lambda))d\lambda}
\end{aligned}
\end{equation}

where $C_x$ and $C_y$ represent the zero-point magnitudes for the target filter
and the observation filter respectively.

at a
blueshifted wavelength given by $F(\lambda / (1+z))$. In other words, it calculates what the observation filter would measure if it was blue shifted.

if the observation filter were blueshifted, similar to the top-right panel of Figure \ref{fig:k-example}, the denominator

the flux that would be measured

The next major advance in K-corrections was

@@

\begin{minipage}[c]{0.95\textwidth}
\lstinputlisting[language=Python, caption=A demonstration that K-corrections
calculated by snpy are off by \texttt{2.5*log10(1+z)}. The roundoff error is
roughly \texttt{2e-16}.]{code_listing.py}
\end{minipage}
